---
title: "New RMD"
output: html_document
date: "2025-02-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(ggplot2)
library(readr)
nyt_bestseller <- read_csv(file.choose("NYT_best_seller .csv"))

# Convert 'bestsellers_date' and 'published_date' to Date format
nyt_bestseller$bestsellers_date <- as.Date(nyt_bestseller$bestsellers_date, format="%m/%d/%y")
nyt_bestseller$published_date <- as.Date(nyt_bestseller$published_date, format="%m/%d/%y")

# Filter for books published between 2010 and 2016
nyt_bestseller <- subset(nyt_bestseller, format(published_date, "%Y") >= 2010 & format(published_date, "%Y") <= 2016)

# Calculate time difference in days between publication and best seller date
nyt_bestseller$days_to_best_seller <- as.numeric(difftime(nyt_bestseller$bestsellers_date, nyt_bestseller$published_date, units="days"))

# Create improved scatter plot with trend line
ggplot(nyt_bestseller, aes(x = published_date, y = days_to_best_seller)) +
  geom_point(aes(color = days_to_best_seller), alpha = 0.6, size = 2) +  # Add color gradient based on days
  geom_smooth(method = "loess", se = TRUE, color = "red", size = 1) +   
  scale_color_gradient(low = "blue", high = "red") +                     # Gradient from blue (low) to red (high)
  scale_x_date(labels = date_format("%Y"), breaks = "1 year") +          # Format x-axis to show months
  labs(title = "Days to Bestseller vs. Publication Date (2010-2016)",
       x = "Publication Date",
       y = "Days to Become Bestseller",
       color = "Days to Bestseller") +
  theme_minimal(base_size = 14) +                                        # Clean theme with larger text
  theme(axis.text.x = element_text(angle = 45, hjust = 1))               # Rotate x-axis labels for readability

```
This data set only had three categories: Trade Fiction, Mass Market, and Hardcover Fiction. This first graph looks at best sellers published from years 2010-2016, and plots the days until days the books reach bestseller list. The range between publication date and days to bestseller is 14 to 15 days.

# Added month for readability 
```{r}

ggplot(nyt_bestseller, aes(x = published_date, y = days_to_best_seller)) +
  geom_point(aes(color = days_to_best_seller), alpha = 0.6, size = 2) +  
  geom_smooth(method = "loess", se = TRUE, color = "red", size = 1) +   
  scale_color_gradient(low = "blue", high = "red") +                     
  scale_x_date(labels = date_format("%b %Y"), breaks = date_breaks("3 months")) +  # show abbreviated months and years
  labs(title = "Days to Bestseller vs. Publication Date (2010-2016)",
       x = "Publication Date",
       y = "Days to Become Bestseller",
       color = "Days to Bestseller") +
  theme_minimal(base_size = 14) +                                        
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
library(dplyr)
print(nyt_bestseller$days_to_best_seller)
```


# Top 10 publishers - Using Alexis' data (Filtered for 2010-2016)

```{r}

# Load required libraries
library(dplyr)


# Read the dataset
nyt_bestseller <- read_csv(file.choose("NYT_best_seller .csv"))

# Convert 'published_date' to Date format
nyt_bestseller <- nyt_bestseller %>%
  mutate(published_date = as.Date(published_date, format="%m/%d/%y"))

# Check date range before filtering
print(range(nyt_bestseller$published_date, na.rm = TRUE))

# Filter data to only include books published between 2010 and 2016
df_filtered <- nyt_bestseller %>%
  filter(published_date >= as.Date("2010-01-01") & published_date <= as.Date("2016-12-31"))

# Count the number of books per publisher
publisher_counts_2010_2016 <- df_filtered %>%
  count(publisher, name = "n") %>%
  arrange(desc(n))

# Ensure data exists before saving
print(dim(publisher_counts_2010_2016))
print(head(publisher_counts_2010_2016))
```
# Visualization for Publisher counts in NYT Best Sellers - Alexis' Data (2010-2016)

```{r echo=TRUE, fig.height=30, fig.width=15}

ggplot(publisher_counts_2010_2016, aes(x = reorder(publisher, n), y = n)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  coord_flip() + 
  labs(title = "Publisher Counts in NYT Best Sellers", 
       subtitle = "Data from 2010-2016",
       x = "Publisher",
       y = "Number of Best Sellers") +
  theme_minimal() +
    theme(
    plot.title = element_text(size = 18, face = "bold"),  # Increase title size
    axis.title.x = element_text(size = 16),  # Increase x-axis title size
    axis.title.y = element_text(size = 16),  # Increase y-axis title size
    axis.text.y = element_text(size = 14),   # Increase y-axis text size (publisher names)
    axis.text.x = element_text(size = 14)    # Increase x-axis text size
  )
```




# Top 10 Publishers 2010-2016

```{r}

top_publishers <- publisher_counts_2010_2016 %>%
  arrange(desc(n)) %>%
  head(10)  # Keep only the top 10 publishers

# Create the ggplot visualization
ggplot(top_publishers, aes(x = reorder(publisher, n), y = n)) + 
  geom_bar(stat = "identity", fill = "red") + 
  coord_flip() + 
  labs(title = "Top 10 Publishers in NYT Best Sellers", 
       subtitle = "Data from 2010-2016",
       x = "Publisher",
       y = "Number of Best Sellers") +
  theme_minimal()

```


Grand Central, Bantam, Berkley, Vintage, Ballantine, Putnam, Dell, Little, Brown, Simon & Schuster, Knopf have most NYT best sellers from 2010-2016 

# Month of publication for best sellers (not looking at categories), using Lucas' data. 


```{r}

# Load the dataset
data <- read_csv(file.choose("dataclean_updated.csv"))

# Convert published_date to Date format
data$published_date <- as.Date(data$published_date, format = "%m/%d/%y")

# Extract month name from published_date
data$month <- format(data$published_date, "%B")

# Order months 
data$month <- factor(data$month, levels = month.name, ordered = TRUE)

# Count number of books published per month
monthly_counts <- data %>%
  group_by(month) %>%
  summarise(count = n())

# Create bar chart
ggplot(monthly_counts, aes(x = month, y = count, fill = month)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "Frequency of Bestseller Books Released Per Month",
       x = "Month",
       y = "Number of Bestsellers") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



# Stacked Area Chart of categories by month with clean data

```{r}


# Convert date column and extract month
data$published_date <- as.Date(data$published_date, format = "%m/%d/%y")
data$month <- format(data$published_date, "%B")

# Order months correctly
data$month <- factor(data$month, levels = month.name)

# Count occurrences of each category per month
area_chart_data <- data %>%
  count(month, New_Category)

# Create stacked area chart
ggplot(area_chart_data, aes(x = month, y = n, fill = New_Category, group = New_Category)) +
  geom_area(position = "stack", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Stacked Area Chart of Bestsellers by Month",
       x = "Month",
       y = "Count of Bestsellers",
       fill = "Category") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Fiction, Manga/Graphic Novels, and Nonfiction categories have the most books on the bestseller list year round compared to the three other categories. 

# Heat map of Bestsellers by month and category with clean data

```{r}

# Convert date column and extract month
data$published_date <- as.Date(data$published_date, format = "%m/%d/%y")
data$month <- format(data$published_date, "%B")

# Order months correctly
data$month <- factor(data$month, levels = month.name)

# Count occurrences of each category per month
heatmap_data <- data %>%
  count(month, New_Category)

# Create heatmap
ggplot(heatmap_data, aes(x = month, y = New_Category, fill = n)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  theme_minimal() +
  labs(title = "Heatmap of Bestsellers by Month and Category", 
       x = "Month", 
       y = "Category", 
       fill = "Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

There is little seasonality for books categorized as Self-Improvement and Other. 
Suprisingly, books categorized as Young Audiences peak on the bestseller list in October. Fiction books peak in March, April, and September and are most frequently on the bestseller list year round compared to other categories. 

# Distribution of bestellers by month and category with clean data

```{r}


# Convert the published_date column to Date format and extract the month
data$published_date <- as.Date(data$published_date, format = "%m/%d/%y")
data$month <- format(data$published_date, "%B")  

# Order months correctly
data$month <- factor(data$month, levels = month.name)

# Create the visualization
ggplot(data, aes(x = month, fill = New_Category)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Distribution of Bestsellers by Month",
       x = "Month",
       y = "Count of Bestsellers",
       fill = "Category") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
``` 

Show distribution of each category per month. Can see that Younger Audiences, Self-Improvement, and Other have least amount of bestsellers year round. Limited seasonality in all categories. 


# Count for each author on best seller list - Using Lucas' Data (CLEANED)

```{r}
nyt_authors <- read.csv(file.choose("dataclean_updated.csv"))

# Count occurrences of each author 
author_counts <- nyt_authors %>% 
  group_by(author) %>% 
  summarise(count = n ()) %>% 
  arrange(desc(count)) 

print(author_counts)
``` 

# Author and their respective catgeory

```{r}


# Count occurrences of each author within each category
author_counts <- nyt_authors %>% 
  group_by(author, New_Category) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  arrange(desc(count))

# Convert to tibble for better display
author_tibble <- as_tibble(author_counts)

# Print the tibble
print(author_tibble)

```

# Visualization of top 20 authors overall

```{r}
top_authors <- author_counts %>% top_n(20, count)

ggplot(top_authors, aes(x = reorder(author, count), y = count, fill = count)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  scale_fill_gradient(low = "blue", high = "red") + 
  labs(title = "Top 20 Authors on NYT Bestseller List", 
       x = "Author", 
       y = "Number of Appearances") + 
  theme_minimal(base_size = 14)
``` 


# Top 10 Authors in Each Category (not excluding if multiple authors have same count)

```{r echo=TRUE, fig.height=20, fig.width=15}



# Count occurrences of each author within each new category
author_counts <- nyt_authors %>% 
  group_by(New_Category, author) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  arrange(New_Category, desc(count))

# Select top 20 authors per category
top_authors <- author_counts %>%
  group_by(New_Category) %>%
  slice_max(order_by = count, n = 10)

# Create bar chart with facet wrap
ggplot(top_authors, aes(x = reorder(author, count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # Flip for readability
  scale_fill_gradient(low = "blue", high = "red") +  # Color gradient
  facet_wrap(~ New_Category, scales = "free_y") +  # Separate plots per category
  labs(title = "Top 10 Authors in Each Category",
       x = "Author",
       y = "Number of Appearances",
       fill = "Count") +
  theme_minimal(base_size = 13) +
  theme(axis.text.y = element_text(size = 12)) +
  theme(axis.text.x = element_text(size = 12))

``` 

# Using with_ties = FALSE to ensure exactly 10 authors per category

```{r fig.height=20, fig.width=15}

library(forcats) 


# Count occurrences of each author within each new category
author_counts <- nyt_authors %>% 
  group_by(New_Category, author) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  arrange(New_Category, desc(count))

# Select top 10 authors per category, ensuring exactly 10 per group
top_authors <- author_counts %>%
  group_by(New_Category) %>%
  slice_max(order_by = count, n = 10, with_ties = FALSE) %>%
  ungroup()  # Remove grouping for plotting

# Create bar chart with facet wrap, ensuring correct ordering within facets
ggplot(top_authors, aes(x = forcats::fct_reorder(author, count, .desc = TRUE), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # Flip for readability
  scale_fill_gradient(low = "blue", high = "red") +  # Color gradient
  facet_wrap(~ New_Category, scales = "free_y") +  # Separate plots per category
  labs(title = "Top 10 Authors in Each Category",
       x = "Author",
       y = "Number of Appearances",
       fill = "Count") +
  theme_minimal(base_size = 13) +
  theme(axis.text.y = element_text(size = 12)) +
  theme(axis.text.x = element_text(size = 12))

```

James Patterson and Danielle Steel appears in multiple categories as a top author. 


```{r echo=TRUE, fig.height=20, fig.width=20}



# Count occurrences of each author within each category
author_counts <- nyt_authors %>% 
  group_by(New_Category, author) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  arrange(New_Category, desc(count))

# Identify authors appearing in more than one category and rank by total count
multi_category_authors <- author_counts %>%
  group_by(author) %>%
  summarise(category_count = n(), total_count = sum(count)) %>%
  filter(category_count > 1) %>%
  arrange(desc(total_count)) %>%
  slice_max(order_by = total_count, n = 10)  # Select top 10 multi-category authors


print(multi_category_authors)

```

Top 11 authors (since Christine Feehan and Lee Child both have 54 books in bestseller list) who have bestsellers in more than 1 category. 
